---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# volundr

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) [![Github commit frequency](https://img.shields.io/github/commit-activity/w/asshah4/volundr)](https://github.com/asshah4/volundr/graphs/commit-activity) [![R-CMD-check](https://github.com/asshah4/volundr/workflows/R-CMD-check/badge.svg)](https://github.com/asshah4/volundr/actions)

<!-- badges: end -->

## Installation

This package has not yet been released on CRAN, but can be downloaded from Github.

```{r, eval = FALSE}
remotes::install_github("asshah4/volundr")
```


## Introduction

The package `volundr` is intended to help forge together multiple models to better understand complex relationships in the causal space. 

## Usage

The package is simple to use. First, lets load the basic packages. The `mtcars` dataset will serve as the example, and we will use linear regressions as the primary test.

```{r}
library(volundr)
library(parsnip)
```

