---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# vlndr

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) 
[![Github commit frequency](https://img.shields.io/github/commit-activity/w/asshah4/vlndr)](https://github.com/asshah4/vlndr/graphs/commit-activity) 
[![R-CMD-check](https://github.com/asshah4/vlndr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/asshah4/vlndr/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

## Installation

This package can be downloaded from CRAN or from [Github](https://github.com/asshah4/vlndr) as below

```{r, eval = FALSE}
# CRAN installation
install.packages("vlndr")
# Or remote/developmental version
remotes::install_github("asshah4/vlndr")
```

## Introduction

The package `vlndr` was intended as a way to handle causal- and epidemiology-based modeling by the following principles:

1. Role determination of variables
2. Generativity in formula creation
3. Multiple model management

The name is an acronym for "**V**ariables of **L**arge **N**umbers; **D**idn't **R**egress" It also lexically represents the word **völundr**, the Norse name for Wayland the Smith, in the spirit of *forging* together many models. 

## Usage

The package is simple to use. First, lets load the basic packages. The `mtcars` dataset will serve as the example, and we will use linear regressions as the primary test.

```{r}
library(vlndr)
library(parsnip)
```

